---
title: 'Statistical report: Characteristics and outcomes of patients with chronic heart failure with or without valvular heart disease – from the ESC HFA EORP Heart Failure Long-Term Registry'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    dev: cairo_pdf
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
   - \usepackage{subfig}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

hf3_lt_fu_data_soladis_jan19.sas7bdat (from the folder DATABASE_ANALYZED.ZIP). 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart", scale_down = F)
```

First patient in: `r min(edata$num_dmVisitdt)` and last patient in: `r max(edata$num_dmVisitdt)`. 

The median age (IQR) is `r edata %>% summarise(med = fn(median(num_age, na.rm = TRUE), dig = 1),
                                             q1 = fn(quantile(num_age, na.rm = TRUE, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(num_age, na.rm = TRUE, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r edata %>% count(num_dmgender) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(num_dmgender == "Female") %>%
  pull(perc)`% females.    

## Created variables

EF is from Echo-Doppler, and if this information is missing, Last known EF. 

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

For patients with missing information on the date of hospitalisation or date of hospitalisation prior to inclusion, 
the time to hospitalisation was imputed with half the time to last
follow-up. Further, times to hospitalisation that were larger than time 
to death or last follow-up were set to death or last follow-up. 

Missing data for the covariates included in the models was imputed with multiple 
imputation using mice [@mice] for 10 datasets and 10 iterations. 
Variables included in the imputation model are indicated in Table \ref{tab:tab1}. 
The primary outcome, all-cause death or first re-hospitalization for HF at follow-up, was included as the Nelson-Aalen estimator. Valve disease was not included in the model or imputed. 

\blandscape

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}
```

\elandscape

```{r, child = "./src/baref.Rmd"}
```

\clearpage

## Associations between valve disease and characteristics

Performed using logistic regression (one for combined VHD vs No VHD and one for Single VDH vs No VHD). 

TO DO: Forestplots? One presenting crud, one adjusted. All variables or selected?

```{r, child = "./src/ortab.Rmd"}

```

\clearpage

Further, crude logistic regressions are performed for MR and TR vs No VHD respectively 
but where the other valve is included (so patients with MR can have TR and vice versa) 
for important valve variables. No adjustment is performed. Imputation is not performed. 

TO DO: Forestplots? 

```{r, child = "./src/ortab_valvevars.Rmd"}

```

\clearpage

## Outcomes

The long-term outcome analysis is performed on the long-term outcome population (patients with follow-up data).

Time is from date of visit and censored at death not defined as an event or end of follow-up.  

The outcomes are presented with cumulative incidence curves and Cox proportional hazards regressions were 
used to model the time to first event. Adjustment was performed for variables indicated in Table \ref{tab:tab1} and the selection was based on clinical judgement. 

The median (q1-q3) follow-up for the long-term outcomes is 
`r edata %>% filter(survpop) %>% summarise(med = fn(median(outtime_death / 365.25 * 12), dig = 1),
                                             q1 = fn(quantile(outtime_death / 365.25 * 12, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(outtime_death / 365.25 * 12, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r edata %>% filter(survpop) %>% 
                                   summarise(sumpy = fn(sum(outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.

```{r, child = "./src/km.Rmd"}

```

\clearpage

```{r, child = "./src/outtab.Rmd"}

```

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) and ==> XXX.

\clearpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/escvalve. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
